### 2016-09-17 mcdonald_nfb_datasharing_endmatter.tex

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
RCC, MPM, SJC, FXC and BL designed the experiment, RCC and SLC developed the real-time DMN tracking system, SLC and JL provided the custom real-time fMRI sequence, ARM, NTVD, JM, and RCC wrote the manuscript, ARM, JM, CF, SG and RCC performed data analyses, JP, ARM, BP, and CF organized data for release, RCC, CCCB, ARM, NTVD, and MB designed and implemented the behavioral tasks and assessments, RT, AMB, RCC, MPM and SJC managed data collection, AA, MMB, ARM, SC, TPC, CG, AG, JG, SH, MK, AL, AMB, LP, HR, CS, ES, RS, MS, ET, KT, BV, LW, and ABW were involved in participant recruitment and performed data collection, VDC, MK, DL, BM, RW, and DW managed data sharing through the COINs pipeline.

\section*{Acknowledgements}
We would like to thank Cathy Hu and Raj Sangoi for MRI operation. Data collection and salary support were provided by NIMH BRAINS R01MH101555.