### 2016-09-17 mcdonald_nfb_datasharing_frontmatter_submit.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Real-time fMRI Neurofeedback Based Stratification of Default Network Regulation Neuroimaging Data Repository.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[aff1]{Amalia R McDonald}
\author[aff1]{Jordan Muraskin}
\author[aff1,aff2]{Nicholas T Van Dam}
\author[aff1]{Caroline Froehlich}
\author[aff1]{Benjamin Puccio}
\author[aff2]{John Pellman}
\author[aff3]{Clemens CC Bauer}
\author[aff1]{Alexis Akeyson}
\author[aff1]{Melissa M Breland}
\author[aff4,aff5]{Vince D Calhoun}
\author[aff1]{Steven Carter}
\author[aff1]{Tiffany P Chang}
\author[aff1]{Chelsea Gessner}
\author[aff1]{Alyssa Gianonne}
\author[aff2]{Steven Giavasis}
\author[aff1]{Jamie Glass}
\author[aff1]{Steven Homan}
\author[aff4]{Margaret King}
\author[aff1]{Melissa Kramer}
\author[aff4]{Drew Landis}
\author[aff1]{Alexis Lieval}
\author[aff6]{Jonathan Lisinski}
\author[aff1,aff7]{Anna Mackay-Brandt}
\author[aff4]{Brittny Miller}
\author[aff1]{Laura Panek}
\author[aff1]{Hayley Reed}
\author[aff1]{Christine Santiago}
\author[aff1]{Eszter Schoell}
\author[aff1]{Richard Sinnig}
\author[aff1]{Melissa Sital}
\author[aff1]{Elise Taverna}
\author[aff1]{Russell Tobe}
\author[aff1]{Kristin Trautman}
\author[aff1]{Betty Varghese}
\author[aff1]{Lauren Walden}
\author[aff4]{Runtang Wang}
\author[aff1]{Abigail B Waters}
\author[aff4]{Dylan Wood}
\author[aff1,aff8]{F Xavier Castellanos}
\author[aff9]{Bennett Leventhal}
\author[aff1]{Stanley J Colcombe}
\author[aff6,aff10,aff11]{Stephen LaConte}
\author[aff1,aff2]{Michael P Milham}
\author[aff1,aff2]{{R Cameron} Craddock\corref{corresp}}
\cortext[corresp]{Corresponding author}
\ead{ccraddock@nki.rfmh.org}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[aff1]{Nathan Kline Institute for Psychiatric Research, Orangeburg, NY, USA }
\address[aff2]{Child Mind Institute, New York, NY, USA }
\address[aff3]{Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Boston, MA, USA } 
\address[aff4]{The Mind Research Network, Albuquerque, New Mexico, USA }
\address[aff5]{Department of Electrical and Computer Engineering, The University of New Mexico, Albuquerque, New Mexico, USA }
\address[aff6]{Virginia Tech Carilion Research Institute, Roanoke, VA, USA }
\address[aff7]{Columbia University, Cognitive Neuroscience Division, Taub Institute and GH Sergeivesky Center, New York, NY, USA }  
\address[aff8]{The Child Study Center, NYU Langone Medical Center, New York, NY, USA } 
\address[aff9]{Department of Psychiatry, University of California – San Francisco, San Francisco, CA, USA }
\address[aff10]{School of Biomedical Engineering and Sciences, Virginia Polytechnic Institute and State University, Blacksburg, VA, USA }  
\address[aff11]{Departments of Emergency Medicine and Emergency Radiology, Virginia Tech Carilion School of Medicine, Roanoke, VA, USA }